---
title: "FinalProject Lipidomics MTBLS352"
author: "Ana Ramirez"
date: "`r Sys.Date()`"
output: html_document
---

## INTRODUCTION
For this document we are using the metabolomics study MTBLS352, available online <https://www.ebi.ac.uk/metabolights/editor/MTBLS352/files> which contains plasma lipidomics data from 293 adults in China, including individuals with normal glucose tolerance, prediabetes, and type 2 diabetes. Quantitative intensity values were not present in the public processed tables; therefore, results are limited to sample counts and annotated lipid identities. Filenames were simplified locally for reproducibility and clarity.

```{r, setup, echo=FALSE, message=FALSE, warning=FALSE}
library(here)
library(tidyverse)
library(janitor)
library(gtsummary)
library(gt)

here::i_am("FinalProject.Rmd")
```


```{r, load-data, echo=FALSE, message=FALSE, warning=FALSE}
data <- readRDS(
  here::here("data", "derived_data", "mtbls352_long.rds"))
```

## Table 1. Lipids identified and number of participants per glycemic group.
Although the study included 293 participants, each individual was measured in multiple assay modes (typically positive and negative ionization). Therefore, the dataset contains 586 sample measurements (assays), and Table 1 summarizes counts at the measurement level rather than unique participants.
```{r, table1, echo=FALSE, message=FALSE, warning=FALSE}
library(here)
table1_df <- readRDS(
  here::here("output", "table1_df.rds"))
knitr::kable(table1_df)
```


## Figure 1. Number of patients per glycemic group
```{r, fig1_png, echo=FALSE, message=FALSE, warning=FALSE}
knitr::include_graphics(
  here::here("figures", "group_barplot.png"))
```
